The Array of Long-Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-06-12 albatros_paper.tex
HCC cleaning up correlator section

% Write text here about the overall final configuration.  Describe the
% site, show synthesized beam, discuss high level system requirements
% (power, data storage, etc).  Include deployment schedule and introduce
% 2-element pathfinder in the following section.

@@

While the basics of 1-bit correlation are well known in the limit where the signal level is much lower than the noise, ALBATROS may well be in the high-signal regime.  We present the basics of 1-bit correlation here, and show that even in the high-signal limit, 1-bit correlation remains viable.
The fundamental output of a 1-bit correlator for real data is $x_{ij} \equiv \left < \tilde{E}_i \tilde{E}_j \right >$.  $ \tilde{E}_i$ is the quantized version of the underlying electric field $E_i$ where $ \tilde{E}_i=1$ for $E_i>0$ and $ \tilde{E}_i=-1$ for $E_i<0$.  Note that complex data can be handled as the combination of real components.  This output is non-linear in the underlying true signal and noise levels, since the output saturates at unity for perfectly correlated electric fields.  To get at the true sky signals, we will need to undo this nonlinearity (the so-called Van Vleck corrections).  For a 2-level correlator, the expected output can be related to the true signals as follows (Van Vleck \& Middleton, e.g. D'Addario):
\begin{eqnarray}
\label{eqn:1bit_output}
\left < \sin(\frac{\pi}{2}x_{ij})\right > = \frac{V_{ij}}{\sqrt{V_{ij}+N_i}\sqrt{V_{ij}+N_j}}
\end{eqnarray}
where $V_{ij}$ is the true visibility, and $V_{ij}+N_{i,j}$ is the total noise power measured by antennas $i$ and $j$.  Inverting this relation to get the true signal requires knowing the noise powers, which is not possible from the 1-bit data themselves.  However, the SNAP board calculates these power levels on timescales of a few seconds, much faster than the power levels change, and so by combining the single-station power measurements with the cross-station cross correlation, we will be able to derive the true sky visibilities.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/corr_efficiency.png}
    \caption{Correlator efficiency $\epsilon_{corr}$ for a 1-bit correlator as a function of signal-to-noise ratio (SNR).  To get the expected SNR from a 1-bit correlator, multiply the SNR derived from the radiometer equation by $\epsilon_{corr}$.  For low SNR, the efficiency is $2/\pi \sim 0.64$, and it drops monotonically as the SNR increases.  The drop is relatively gentle, only changing by a factor of $\sim 2$ for SNR=2.  For baseline separations of many wavelengths, the correlated component of the electric field is likely to be a small fraction of the total power, so ALBATROS is unlikely to lose more than a factor of $\sim 2$ by using 1-bit correlation.}
    \label{fig:1bit_efficiency}
    \end{center}
\end{figure}

@@

>>>>

@@

The autonomous station can be configured to record 1-, 2-, or 4-bit
baseband.  Signal fidelity improves with higher numbers of bits, but
the penalty is increased data volume.  Here we discuss the prospects
of employing 1-bit correlation for the final \albatros\ array to keep
the baseband data volume to a minimum.  While the basics of 1-bit
correlation are well known in the limit where the signal level is much
lower than the noise, \albatros\ may well be in the high-signal
regime.  The fundamental output of a 1-bit correlator for real data is
\begin{equation}
  x_{ij} \equiv \left < \tilde{E}_i \tilde{E}_j \right >.
\end{equation}
Here $\tilde{E}_i$ is the quantized version of the underlying electric
field $E_i$:
\begin{equation}
  \tilde{E}_i=
  \begin{cases}
    1 & \text{if}\ E_i > 0 \\
    -1 & \text{if}\ E_i < 0.
  \end{cases}
\end{equation}
Complex data can be handled as the combination of real components.
This output is non-linear in the underlying true signal and noise
levels, since the output saturates at unity for perfectly correlated
electric fields.  To get at the true sky signals, we will need to undo
this nonlinearity (the so-called Van Vleck corrections).  For a
2-level correlator, the expected output can be related to the true
signals as follows (Van Vleck \& Middleton, e.g. D'Addario):
\begin{eqnarray}
\label{eqn:1bit_output}
\left < \sin(\frac{\pi}{2}x_{ij})\right > = \frac{V_{ij}}{\sqrt{V_{ij}+N_i}\sqrt{V_{ij}+N_j}}
\end{eqnarray}
where $V_{ij}$ is the true visibility, and $V_{ij}+N_{i,j}$ is the total noise power measured by antennas $i$ and $j$.  Inverting this relation to get the true signal requires knowing the noise powers, which is not possible from the 1-bit data themselves.  However, the SNAP board calculates these power levels on timescales of a few seconds, much faster than the power levels change, and so by combining the single-station power measurements with the cross-station cross correlation, we will be able to derive the true sky visibilities.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/corr_efficiency/corr_efficiency.png}
    \caption{Correlator efficiency $\epsilon_{corr}$ for a 1-bit correlator as a function of signal-to-noise ratio (SNR).  To get the expected SNR from a 1-bit correlator, multiply the SNR derived from the radiometer equation by $\epsilon_{corr}$.  For low SNR, the efficiency is $2/\pi \sim 0.64$, and it drops monotonically as the SNR increases.  The drop is relatively gentle, only changing by a factor of $\sim 2$ for SNR=2.  For baseline separations of many wavelengths, the correlated component of the electric field is likely to be a small fraction of the total power, so ALBATROS is unlikely to lose more than a factor of $\sim 2$ by using 1-bit correlation.}
    \label{fig:1bit_efficiency}
    \end{center}
\end{figure}

@@

% We present the basics of 1-bit correlation here, and show that even in
% the high-signal limit, 1-bit correlation remains viable.
